---
title: Graphics and maps with R
subtitle: Under Hadley Wickham's reign
author: Marie-Pierre Etienne
date: "2020/09/11 (updated: `r Sys.Date()`)"
institute: "https://github.com/MarieEtienne/stateoftheRinRennes/"
csl: "../resources/apa-no-doi-no-issue.csl"
output:
  beamer_presentation:
    slide_level: 3
    highlight: kate
    keep_tex: true
    includes:
      in_header: ../resources/common_preamble.tex
      before_body: ../resources/common_startup.tex
bibliography: visu.bib
fontsize: 10pt
params: 
  child_path : ""
  setup_path : "../resources/"
---

```{r setup, include = FALSE, cahe = FALSE}
source(paste0(params$setup_path, "knitr_setup.R"))
with_sol <- TRUE ## in order to control the output
with_course <- TRUE
```


```{r, load_refs, include=FALSE, cache=FALSE}
library(kableExtra)
## textformating function
source('../utils/text_formatting.R')
```

#  Introduction


###  Tutorial based on 
- [ggplot2: Elegant Graphics for Data Analysis](https://ggplot2-book.org/index.html) @wickham2016ggplot2
- [Geocomputation with R](https://geocompr.robinlovelace.net/) @lovelace2019geocomputation`
- [Tutoriel de Raphaelle Momal et Marie Perrot Dockès](https://stateofther.github.io/post/fancy-plotting/)
- [Drawing beautiful maps programmatically with R](https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html)



### ggplot presentation

gg stands for   *Grammar of Graphics*  and  ggplot2  is *based on the Grammar of Graphics @wilkinson2005grammar`, that allows you to compose graphs by combining independent components* (@wickham2016ggplot2).

*[The] grammar tells us that a graphic maps `r colFmt("the data", "darj1")` to the `r colFmt("aesthetic attributes", "darj2")` (colour, shape, size) of `r colFmt("geometric objects",  "darj4")`  (points, lines, bars). The plot may also include statistical transformations of the data and information about the plot’s coordinate system. Facetting can be used to plot for different subsets of the data. The combination of these independent components are what make up a graphic.*  (@wickham2016ggplot2). -->


For coherence, we will make use of the `tidyverse` suite.

```{r load_tidyverse, message = FALSE}
library(tidyverse)
```



### First running example:  The Palmer Penguins dataset 

The Palmer penguins 
```{r palmer_data}
#remotes::install_github("allisonhorst/palmerpenguins")
data(penguins,package = "palmerpenguins")
penguins <- penguins %>% 
  rename(bill_l = bill_length_mm, bill_d = bill_depth_mm, flip_l = flipper_length_mm, bm = body_mass_g) 
penguins %>% 
  print(n=2)

```

### The Palmer Penguins dataset : overview


```{r show_data_palmer}

penguins %>% 
  group_by(species, sex, year, island) %>% 
  mutate(n = n()) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE) %>% 
  print(n=10)
```


# Simple plots

## Scatter plot

### Base information

```{r scatter_plot_void}
ggplot( data = penguins) + aes( x= bill_l, y=bill_d) 

```

### First scatter plot 

```{r scatter_plot_base}
ggplot( data = penguins) + aes( x= bill_l, y=bill_d) + geom_point()

```




### Adding colors

```{r scatter_plot1_species}
gg <-  ggplot( data = penguins) + 
  aes( x= bill_l, y=bill_d, col = species) + 
  geom_point()
gg
```

###  Color blind compliant palette

```{r scatter_plot1_viridis}
gg + scale_color_viridis_d()
```

### Wesanderson palette

Following  a Tumblr blog  [Wes Anderson Palettes](https://wesandersonpalettes.tumblr.com/), Karthik Ram proposes the [wesanderson palette](https://github.com/karthik/wesanderson) on Github. 


```{r scatter_wesanderson}
## remotes::install_github("wesanderson") 
color_darj <- wesanderson::wes_palette(name = "Darjeeling1")
gg + scale_color_manual(values = color_darj)
```

### Changing  labels

```{r scatter_labels}
gg <- gg + scale_color_manual(values = color_darj) +
  labs( x = 'Bill length in mm',  y = 'Bill depth in mm',  color = "Species")
gg
```

### Pre set general theme

```{r scatter_themelight}
gg <- gg + scale_color_manual(values = color_darj) +
  labs( x = 'Bill length in mm',  y = 'Bill depth in mm',  color = "Species") +
  theme_light()
gg
```

### Pre set general theme

```{r scatter_thememinimal}
gg + scale_color_manual(values = color_darj) +
  labs( x = 'Bill length in mm',  y = 'Bill depth in mm',  color = "Species") +
  theme_minimal()
```


### Changing legend position

The default theme might not be the best option 

```{r scatter_below}
gg + theme(legend.position="bottom")
```

### Customed legend position 

The legend position can also be specified by a vector ` c(x,y)`.  Their values should be between 0 and 1. c(0,0) corresponds to the “bottom left” and c(1,1) corresponds to the “top right” position. We may want to change the size.

```{r scatter_within}
library(extrafont)
loadfonts()
gg + theme(legend.position=c(.9, .6))
```

### Customizde theme

```{r scatter_custom}
## install.packages('extrafont')
library(extrafont)
## font_import() # just run once after installation
loadfonts()
gg + theme(legend.position=c(.9, .6),
           text = element_text(size = 8, face = "italic"),
           axis.text.x = element_text(angle=90, hjust=1), 
           legend.text = element_text(size = 9, face = 'plain'),
           legend.title = element_text(size = 10, face = "bold") )

```


### Adding statistical transformation

Adding a fitted linear regression 

```{r stat_info_lm}
gg + geom_smooth(method = 'lm', se = FALSE)
```


### Adding statistical transformation

Adding a fitted loess regression  with standaer estimation error 

```{r stat_info}
gg + geom_smooth(method = 'loess', se = TRUE)
```



## Other geometries

### Histogram

```{r simple_hist}

penguins %>% 
  ggplot() + aes(x = bill_l) + geom_histogram() +
  labs( x = 'Bill length in mm') +
  theme_minimal()

```


### Histogram - exercise

Add the color according to species and use the `GrandBudapest1` palette from `wesanderson` package.

```{r color_hist, eval = FALSE, echo = TRUE}

penguins %>%  ...
```



### Histogram - exercise

```{r color_hist_cor}

penguins %>%
  ggplot() + aes(x = bill_l, fill = species) + geom_histogram() +
  labs( x = 'Bill length in mm') +  
  scale_fill_manual(values = wesanderson::wes_palette('GrandBudapest1', n = 3)) +
  theme_minimal()
```


### Split in 3 histograms

```{r color_hist_facet}

penguins %>%
  ggplot() + aes(x = bill_l, fill = species) + 
  facet_wrap(~species) + geom_histogram() +
  labs( x = 'Bill length in mm') +  
  scale_fill_manual(values = wesanderson::wes_palette('Darjeeling1')) +
  theme_minimal()
```

### Split in 3 histograms  with fitted densities

Change the previous graph to add a fitted density line

```{r color_hist_dens_ex, eval = FALSE}

penguins %>% ...
```


### Split in 3 histograms  with fitted densities

Change the previous graph to add a fitted density line

```{r color_hist_dens_cor}

 penguins %>%  ggplot() + aes(x = bill_l, y = ..density..) + 
  facet_wrap(~species) + geom_histogram(alpha=0.5, aes( fill = species)) +
   geom_density(aes(col = species)) +
  labs( x = 'Bill length in mm') +  
  scale_fill_manual(values = wesanderson::wes_palette('Darjeeling1')) +
  scale_color_manual(values = wesanderson::wes_palette('Darjeeling1')) +
  theme_minimal()
```



## Useful and/or fancy packages


### For publications

The `ggpubr' package is very helpful for publication.

```{r ggpubr_p1}
gg_p1 <- gg
gg_p1
```

### For publications

The `ggpubr' package is very helpful for publication.

```{r ggpubr_p2}
gg_p2 <- penguins %>%
  ggplot()  + aes(x = bill_l, y = ..density..) + geom_histogram(alpha=0.5, aes( fill = species)) +
   geom_density(aes(col = species)) +
  labs( x = 'Bill length in mm') +  
  scale_fill_manual(values = wesanderson::wes_palette('Darjeeling1')) +
  scale_color_manual(values = wesanderson::wes_palette('Darjeeling1')) +
  theme_minimal()
gg_p2
```


### For publications

The `ggpubr' package is very helpful for publication.

```{r ggpubr_p3}
##install.packages('ggpubr')  
ggpubr::ggarrange(gg_p1, gg_p2, nrow=2, ncol = 1)
```

### For publications

The `ggpubr' package is very helpful for publication.

```{r ggpubr_p4}
ggpubr::ggarrange(gg_p1, gg_p2, nrow=2, ncol = 1, common.legend = TRUE)
```

### For bivaraite exploration
```{r ggpairs}
##install.packages('GGally')
library(GGally)
penguins %>% ggpairs(columns = c(1,3,4,5), mapping = aes(col = species)) + 
  scale_color_manual(values = wesanderson::wes_palette('Darjeeling1'))+ 
  scale_fill_manual(values = wesanderson::wes_palette('Darjeeling1')) + theme(text = element_text(size = 6))
```


### For bivaraite exploration and nice looking options
```{r ggpairs_fancy}
penguins %>% ggpairs(columns = c(1,3,4,5), mapping = aes(col = species),
                     upper = list(continuous = wrap( "cor",size = 2)),
                     lower = list(continuous = wrap('points', size = .5))) + 
  scale_color_manual(values = wesanderson::wes_palette('Darjeeling1'))+ 
  scale_fill_manual(values = wesanderson::wes_palette('Darjeeling1')) + theme(text = element_text(size = 6))
```


### For presentation - `gganimate`



```{r ggpairs_fancy}
penguins %>% ggpairs(columns = c(1,3,4,5), mapping = aes(col = species),
                     upper = list(continuous = wrap( "cor",size = 2)),
                     lower = list(continuous = wrap('points', size = .5))) + 
  scale_color_manual(values = wesanderson::wes_palette('Darjeeling1'))+ 
  scale_fill_manual(values = wesanderson::wes_palette('Darjeeling1')) + theme(text = element_text(size = 6))
```


# References
